---
title: "SubgroupBoost tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SubgroupBoost}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SubgroupBoost)
```

## Introduction

This tutorial guide the reader through the analysis of treatment performing and non-performing subgroup identification using our package SubgroupBoost. 

## Data input and preprocessing

* Load data. In this tutorial, we will use a simulated data that comes with SubgroupBoost package to demonstrate. 

```{r data}
data("simdata")
names(simdata)
str(simdata$train)
str(simdata$test)
```

* The data is a list containing 4 elements: training data, testing data, true subgroup label for training data, and true subgroup label for testing data. The training data contains 600 samples, and testing data contains 5000 samples. The data columns has 26 variables: `trt01p` is treatment indicator. `evnt1` and `aval1` are for the primary outcome overall survival (OS). `evnt1` is the indicator of event (death) for OS, and `aval1` is the time value. `evnt2` and `aval2` are for the secondary outcome time-to-progression (TTP), where `evnt2` is the indicator of event (disease progression), and `aval2` is the time value. Variable `s1` is the predictive variable that interact with treatment, while variables z1 to z20 are not related to treatment effect.

* Preprocess the training and testing data
```{r}
#----- data manipulation-----#
#training data
dat.train=simdata[[1]][,-c(1:5)]
info.train=simdata[[1]][,1:5]
dtrain <- xgb.DMatrix(as.matrix(dat.train),label = info.train$trt01p)

#testing data
dat.test=simdata[[2]][,-c(1:5)]
info.test=simdata[[2]][,1:5]
dtest <- xgb.DMatrix(as.matrix(dat.test),label = info.test$trt01p)
```

## RMST
* Use RMST based gradient tree boosting method to identify subgroups of patients. Here we choose OS as outcome, as RMST based method can only take one survival outcome.
```{r RMST}
#----- RMST death -----#
dat1=data.frame(dat.train,trt01p=info.train$trt01p, evnt=info.train$evnt1, aval=info.train$aval1)
set.seed(123) # for reproducibility
model_RMST<-SubgroupBoost.RMST(dat1)

#check variables selected
importance=xgb.importance(model_RMST$feature_names, model_RMST)$Feature
importance

# Get predicted value for each sample in training data
pred_RMST<-predict(model_RMST, dtrain)
# Calcualate probabilities of belonging to treatment performing subgroup by sigmoid transformation
prob_RMST<-1/(1+exp(-pred_RMST))
#plot the resulting probabilities
hist(prob_RMST)

#Wrap it up to be a function, use probability 0.5 as cutoff for subgroup label
SubgroupBoost.predict<-function(model, data){
  # Get predicted value of model
  pred<-predict(model, data)
  # Calcualate probabilities by sigmoid transformation
  prob<-1/(1+exp(-pred))
  # Use 0.5 cutoff to seperate samples into two subgroups
  assign<-ifelse(prob>0.5,1,0)
  return(group_assign=assign)
}

#2 by 2 table of predicted and true subgroup label for training and testing data
RMST.predict.train<-SubgroupBoost.predict(model_RMST, dtrain)
table(Predicted=RMST.predict.train, True=simdata[[3]])

RMST.predict.test<-SubgroupBoost.predict(model_RMST, dtest)
table(Predicted=RMST.predict.test, True=simdata[[4]])
```

##Win-difference
* Use win-difference based gradient tree boosting method to identify subgroups of patients. Here we take both OS and TTP as outcomes for win-ratio, OS surives as the primary endpoint and TTP as the secondary outcome, i.e., surrogate of OS when OS is censored. Win-difference method can take one or multiple outcomes, it can be survival, continous or 
```{r Winratio, eval=FALSE}
#----- win-ratio ----#
#Specify the type of outcome by argument comparison. For two survival outcomes, set it equals "survival survival". 
#Note: Make sure the Primary outcome is defined in variable evnt1 and aval1, and secondary outcome is defined by columns evnt2 and aval2 in info.train object, for two survival outcomes case.
comparison="survival survival"
model_wd<-SubgroupBoost.wd(dat.train, info.train, comparison)

#check variables selected
importance=xgb.importance(model_wd$feature_names, model_wd)$Feature
importance

#2 by 2 table of predicted and true subgroup label for training and testing data
wd.predict.train<-SubgroupBoost.predict(model_wd, dtrain)
table(Predicted=wd.predict.train, True=simdata[[3]])

wd.predict.test<-SubgroupBoost.predict(model_wd, dtest)
table(Predicted=wd.predict.test, True=simdata[[4]])

```

